Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread

### 2017-06-29 paper.tex
added new reference to potato line in table
edits to domestication section

\title{Review: Adaptive Introgression has Expanded the Genetic Base of Crops during post-Domestication Spread}

@@

The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation. \mbh{will need citation for this, possibly Gepts review or a Harlan paper}
However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and more diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
An important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
Maize was domesticated from (\emph{Zea mays} ssp. \emph{parviglumis}) approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
%The following is not fully clear.  Cut or revise.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
Height differences were not detected under lowland conditions.
%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.

@@

A framework in which crops were domesticated from a single population or even a single species is, in many instances, an oversimplification.
An history of introgression during geographic expansion increasingly appears to be quite common.
The addition of gene flow to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:

\subsection*{What is the progenitor of a crop?}
When  highly-introgressed or hybrid populations are selected for domestication (as in the case of potato and tomato), identification of original progenitor(s) and domestication centers is difficult.
Determining whether a crop 
Crops arising from highly-interfertile compleces of wild relatives

>>>>

\title{Review: Adaptive Introgression Expanded the Genetic Base of Crops during post-Domestication Spread}

@@

The process of domestication is often conceptualized as geographically constrained, with crops originating from a wild progenitor within one or more defined centers followed by expansion to the modern-day extent of cultivation. \mbh{will need citation for this, possibly Gepts review or a Harlan paper}
However, archaeological and genetic evidence are beginning to reveal that, in many cases, domestication has been temporally protracted and geographically diffuse \cite{brown2009complex, Meyer2016, wang2017, zhou2017, Fuller2014}.
An important aspect of the emerging complexity of domestication is beneficial gene flow (\emph{i.e.}, adaptive introgression) from locally adapted wild relatives during crop expansion following initial domestication.

@@

The relationship between maize (\emph{Zea mays} ssp. \emph{mays}) and the teosinte \emph{Zea mays} ssp. \emph{mexicana} (hereafter referred to as \emph{mexicana}) offers a prime case study of adaptive wild-to-crop introgression.
Maize was domesticated from \emph{Zea mays} ssp. \emph{parviglumis} (hereafter referred to as \emph{parviglumis})  approximately 9,000 years ago in the lowlands of the Balsas River Valley in Mexico \cite{matsuoka2002single}.
From this domestication center, maize spread into the highlands of the Mexican Central Plateau, where it came into sympatry with \emph{mexicana}.
Introgression from \emph{mexicana} to maize in the Central Plateau has long been reported based on both morphological \cite {wilkes1977, lauter2004, doebley1984} and molecular \cite{matsuoka2002, vanHeerwaarden2011, doebley1987, warburton2011, fukunaga2005} data.
However, \citep{hufford2013} first localized \emph{mexicana} introgression to chromosomal regions and provided evidence that it was likely adaptive.
%The following is not fully clear.  Cut or revise.
The authors identified nine genomic regions in several maize populations which consistently showed evidence of \emph{mexicana} introgression based on chromosome painting using both HAPMIX and the linkage model of STRUCTURE (Figure \ref{fig:introgressionMaize}).
These introgressed segments showed low diversity and overlapped QTL that had previously been found to control anthocyanin content and leaf macrohairs \cite{lauter2004}, traits known to be adaptive at high elevation.
In a growth chamber experiment, the authors demonstrated that maize populations with \emph{mexicana} introgression showed greater plant height (a proxy for fitness) under highland environmental conditions than populations that lacked introgression.
Height differences were not detected under lowland conditions.
%Among the nine regions, three span the centromeres of chromosomes 5, 6, and 10, and one is located in the inversion polymorphism on chromosome 4, suggesting a significant role of genome structures restricting recombination in adaptive introgression.
%In summary, the introgressed alleles/haplotypes from \emph{mexicana} to maize conferred adaptation to highland habitats when maize migrated from Mexican lowlands.

@@

A framework in which crops were domesticated from a single population or even a single species is, in many instances, an oversimplification.
A history of introgression during the geographic expansion of crops increasingly appears common.
The addition of ongoing gene flow and adaptive introgression to our understanding of crop demography could bear importantly on fundamental questions of crop domestication:

\subsection*{What is the progenitor of a crop?}
Depending on the extent of post-domestication gene flow with new wild relatives, identification of a crop's progenitor can be complicated or completely confounded.
The level of divergence of a crop from these newly encountered populations will decrease due to introgression, a signal that could be mistaken for origin rather than gene flow.
For example, when determining a single origin of maize from \emph{parviglumis}, Matsuoka and colleagues \cite{Matsuoka2002} identified a paradox: while \emph{parviglumis} is found exclusively in the lowlands of southwest Mexico, maize with allele frequencies most similar to \emph{parviglumis} was found in the highlands of the Mexican Central Plateau.
Several years later, van Heerwaarden \emph{et al.} \cite{vanHeerwaarden2011} resolved the paradox by determining widespread introgression in the highlands from \emph{mexicana}, which is closely related to \emph{parviglumis}, has caused maize from this region to appear ancestral.
Similarly, extensive post-domestication introgression from wild relatives in both potato \cite{Spooner2014} and tomato
